### 2016-04-04 CixPAG/paper58.Rmd

*Benestad et al - Upper-limit estimation of future precipitation return-values* 
-----------------------------------
title: "Analysis and calculations in manuscript"
author: "Rasmus Benestad"
date: "April 4, 2016"
output: pdf_document
---

First pre-amble that checks whether the esd-package is installed and installs it if needed.

```{r}
rm(list=ls())
#xlim <- c(-10,40); ylim <- c(35,70)
xlim <- c(0,35); ylim <- c(45,70)
readecad <- FALSE
figshare=TRUE
nmin=50

## Check if you need to get the esd-package:
install.esd <- ("esd" %in% rownames(installed.packages()) == FALSE)

if (install.esd) {
  print('Need to install the esd package')
  ## Need online access.
  ## Use the devtools-package for simple facilitation of installing.
  if ("devtools" %in% rownames(installed.packages()) == FALSE)
    install.packages('devtools')
  library(devtools)
  ## Install esd directly from github
  install_github('metno/esd')
  print('The latest version of esd has been installed')
}

## Start the esd-library:
library(esd)
```

Below are definition of a number of functions used in the data processing and analysis. The first are simple functions used in ploting.

**Simple

```{r}
## Standardisation
s <- function(x) {
    z <- coredata(x)
    y <- (z - mean(z,na.rm=TRUE))/sd(z,na.rm=TRUE)
    attributes(y) <- NULL
    coredata(x) <- y
    return(x)
}
```

**Functions which estimate various statistics**

```{r}
## Return statistics for mean,min,max,wettest month,driest month
muclim <- function(x) {
  y <- coredata(x)
  iX <- mean((1:12)[is.element(y,max(y))])
  iN <- mean((1:12)[is.element(y,min(y))])
  stats <- c(mean(y,na.rm=TRUE),min(y,na.rm=TRUE),max(y,na.rm=TRUE),iX,iN)
  names(stats) <- c('mean','min','max','wettest month','driest month')
  stats
}

## Produce a set of percentiles and their counterparts for the exponential distribution
qqexp <- function(x,x0=1) {
  x[x < x0] <- NA
  if (sum(is.finite(x))>0) {
    mu <- mean(x,na.rm=TRUE)
    qx <- quantile(x,probs=seq(0,1,length=101),na.rm=TRUE)
    qy <- -log(1-seq(0,1,length=101))*mu
  } else {
    qx <- rep(NA,101); qy <- rep(NA,101)
  }
  return(cbind(qx,qy))
}

## Skill associated with predicting the wet-day mean mu
muskill <- function(x) {
  ## Estimate the skill of the calibration:
  r2 <- round(summary(lm(y ~ x,data=x))$r.squared,3)
  ## Negative slopes are not credible:
  if (summary(lm(y ~ x,data=x))$coefficients[2] < 0) r2 <- 0
  r2
}

## A function for extracting the regression coefficients and their
## error terms.
beta <- function(x,verbose=FALSE) {
  wc.model <- lm(y ~ x, data=x)
  if (verbose) print(summary(wc.model))
  beta <- summary(wc.model)$coefficients[c(2,4)]
  return(beta)
}
```

**Data input and procesing**

## Read the data from the CMIP5 GCMs
```{r}
readGCMs <- function(path='CMIP5.monthly/rcp45/',pattern='tas',
                     lon=c(-100,30),lat=c(0,40)) {
    ncfiles <- list.files(path=path,pattern=pattern,full.names=TRUE)
    n <- length(ncfiles)
    print(paste(n,'netCDF files'))
    X <- matrix(rep(NA,n*201),n,201)
    for (i in 1:n) {      print(paste(i,ncfiles[i]))
      ## Spatial average:                    
      gcm <- annual(spatial.avg.field(C.C.eq(retrieve(ncfiles[i],
                                                      lon=lon,lat=lat))))
      i1 <- is.element(1900:2100,year(gcm))
      i2 <- is.element(year(gcm),1900:2100)
      X[i,i1] <- coredata(gcm)[i2]
    }
    ## Extract the 5 & 95 percentile and the ensemble mean:
    print('Extract the 5 & 95 percentile and the ensemble mean')
    x <- apply(X,2,function(x) c(quantile(x,probs=c(0.05,0.95),na.rm=TRUE),mean(x,na.rm=TRUE)))
    print(dim(x))
    names(x) <- c('q05','q95','mean')
    x <- zoo(t(x),order.by=1900:2100)
    w2000 <- X[,is.element(1900:2100,2000)]
    w2050 <- X[,is.element(1900:2100,2020)]
    w2100 <- X[,is.element(1900:2100,2100)]
    plot(x,plot.type='single')
    attr(x,'path') <- path
    attr(x,'2000') <- w2000
    attr(x,'2050') <- w2050
    attr(x,'2100') <- w2100
    attr(x,'N') <- n
    attr(x,'region') <- paste(lon,lat,collapse=' ')
    return(x)
}
```

**model calibration and regression analysis**

```{r}
## Calibrate a model for the wet-day mean mu using temperature as input
mucal <- function(x,pre=NULL,lon=c(-100,30),lat=c(0,40),
                  plot=FALSE,verbose=FALSE) {

## If no pre, use the crude NCEP-reanalysis provided in esd 
  if (is.null(pre)) {
    if (verbose) print('default predictor')
      t2m <- t2m.NCEP(lon=lon,lat=lat)
      pre <- spatial.avg.field(C.C.eq(t2m))
      if (plot) plot(EOF(t2m))
  } else
      if (is.character(pre))
          pre <- spatial.avg.field(C.C.eq(retrieve(ncfile=pre,lon=lon,lat=lat))) else
  if (inherits(pre,'field')) {
      if (is.T(pre)) pre <- spatial.avg.field(C.C.eq(pre)) else
      pre <- spatial.avg.field(pre)
  } else if (inherits(pre,'station')) pre <- pre
  z <- aggregate(pre,by=month,FUN='mean')

  cal <- data.frame(y=coredata(x),x=coredata(z))
  attr(cal,'standard.error') <- attr(x,'standard.error')
  stats <- cor.test(cal$y,cal$x)
  wc.model <- lm(y ~ x, data=cal)
  if (plot) {
      dev.new()
      par(bty='n',cex.sub=0.7,col.sub='grey40')
      ylim <- range(cal$y,na.rm=TRUE); xlim=range(cal$x,na.rm=TRUE)
      dy <- diff(ylim)/25
      plot(cal$x,cal$y,pch=19,cex=1.5,col='grey',
           ylab=expression(paste(mu,' (mm/day)')),
         xlab=expression(paste(e[s],' (Pa)')),
         ylim=ylim,xlim=xlim,
         main='"Worst-case" fit based on seasonal variations',
         sub=paste(loc(x),' (',round(lon(x),2),'E/',round(lat(x),2),'N; ',
             alt(x),'m.a.s.l.)',sep=''))
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y+2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y-2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x+2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x-2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      points(cal$x,cal$y,pch=19,cex=1.5,col='grey')
      grid()
      abline(wc.model)
      text(xlim[1],ylim[2],paste('Correlation=',round(stats$estimate,2),
                                 '(','p-value=',
                                 100*round(stats$p.value,4),'%)'),
           pos=4,cex=0.7,col='grey')
      text(xlim[1],ylim[2]-dy,paste('Regression: y=',
                                    round(wc.model$coeff[1],4), '+',
                                    round(wc.model$coeff[2],4), 'x (R2=',
                                    round(summary(wc.model)$r.squared,2),')'),
           pos=4,cex=0.7,col='grey')
      par(new=TRUE,fig=c(0.5,0.97,0.1,0.5),yaxt='n',xaxt='n',xpd=TRUE,
          cex.axis=0.7,col.axis='grey')
      plot((cal$x - mean(cal$x))/sd(cal$x),type='l',lwd=2,
           ylab='',xlab='',col=rgb(0.6,0.3,0))
      lines((cal$y - mean(cal$y))/sd(cal$y),type='l',lwd=2,col=rgb(0,0.3,0.6))
      par(xaxt = "s")
      axis(1,at=1:12,labels=month.abb,col='grey')
  }
  invisible(cal)
}
```

Functions for projection and prediction for the future/past

```{r}
## Projection based on the calibration with the annual cycle:
muproject <- function(x,gcm,verbose=FALSE,prct=TRUE) {
  #print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
    print(dim(gcm))
  }
  pq05 <- data.frame(x=coredata(gcm[,1]))
  pq95 <- data.frame(x=coredata(gcm[,2]))
  pmea <- data.frame(x=coredata(gcm[,3]))
  y <- cbind(predict(wcmodel,newdata=pq05),
             predict(wcmodel,newdata=pq95),
             predict(wcmodel,newdata=pmea))
  if (verbose) print(dim(y))
  if (prct) {
#    ii <- is.element(year(gcm),1961:1990)
    ii <- is.element(year(gcm),2000:2010)
    bline <- mean(y[ii,3])
    #bline <- colMeans(y[ii,])
    #browser()
    #print(length(bline))
    #y <- t(100*t(y)/bline)
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(gcm))
  names(y) <- names(gcm)
  return(y)          
}

## Predict values of the wet-day mean mu taking a given predictor
mupredict <- function(x,pre,verbose=FALSE,prct=TRUE) {
  if (verbose) print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
  }
  eval <- data.frame(x=coredata(pre))
  y <- predict(wcmodel,newdata=eval)
  if (prct) {
    ii <- is.element(year(pre),1961:1990)
    bline <- mean(y[ii])
    if (verbose) print(length(bline))
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(pre))
  names(y) <- 'mu'
  return(y)          
}
```

**Generation of graphics/maps**

```{r}
## Generate a map for the PC weights for the different modes of the
## wet-day mean mu annual cycle.
mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
  #dev.new()
  r2 <- as.numeric(r2)
  colpc <- rev(colscal("t2m",n=100))
  cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
  cz[cz > 100] <- 100; cz[cz < 1] <- 1
  #print(c(length(cz),sum(is.finite(cz))))
  col <- colpc[cz]
  pch <- rep(19,length(r2))
  mo <- c(r2) > 0.4
  hi <- c(r2) > 0.6
  pch[!hi] <- 1
  pch[!mo] <- 4
  cex <- 1.5*c(r2) + 0.2
  #print(c(sum(is.finite(pch)),length(col),sum(is.finite(cex))))

  map(mu, xlim=xlim,ylim=ylim,bg='grey80',new=FALSE)
  points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
  #text(lon(mu),lat(mu),r2,cex=0.7); print(table(col))
  par(xpd=TRUE)
  text(10,73,paste('Annual cycle in PC',ipca,
      'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)

  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.05, 0.1, 0.05, 0.2))

  par(xaxt = "n", yaxt = "s", fig = c(0.05,0.25,0.80,0.95),
      mar = c(0, 1, 0, 0), new = TRUE, las = 1, cex.axis = 0.5,bty='n')
  plot(pca$u[,ipca],type='l',lwd=3,col="red")
  par(xaxt = "s")
  axis(1,at=1:12,labels=month.abb)
}

## Plot shaded regions
shade <- function(x,col=rgb(0.5,0.5,0.5,0.3),border=NULL) {
  t <- index(x)
  if (is.null(border)) border <- col
  y <- coredata(x)
  polygon(c(t,rev(t)),c(y[,1],rev(y[,2])),col=col,border=border)
  lines(t,y[,3],lwd=5,col=col)
}

## Estimate correlation between different sets of wet-day mean mu
## The function is designed to be used in apply for best efficiency
cormu <- function(x) {
  n <- length(x); nh <- n/2
  x1 <- x[1:nh]; x2 <- x[(nh+1):n]
  ok <- is.finite(x1) & is.finite(x2)
  return(cor(x1[ok],x2[ok]))
}
```

The definition of functions is followed by code that carry out the analysis based on these.  First get the data needed:

```{r}
##-----------------------------------------------------------------------------------------

## Need to obtain some of the data files - fetch from Figshare:

if (!file.exists("mu.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193033/mu.worstcasemu.rda",destfile="mu.worstcasemu.rda") 
}

if (!file.exists("pre.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193038/pre.worstcasemu.rda",destfile="pre.worstcasemu.rda") 
}

if (!file.exists("cmip5.rda") & figshare) {
  download.file("http://files.figshare.com/2193041/cmip5.rda",destfile="cmip5.rda") 
}

## Preparations that only needs to be done once. 
if (readecad) {
## This section generates a processed data file from scratch using ECA&D data:
  pr <- station(src=c('metnod','ecad'),param='precip',nmin=nmin,it=c(1961,2010),lon=xlim,lat=ylim)
  pr <- subset(pr,it=c(1961,2014))
  nt <- apply(pr,2,FUN='nv')
  pr <- subset(pr,is=(nt >=19000))
  cpr <- coredata(pr)
  cpr[cpr > 250] <- NA
  coredata(pr) <- cpr
  save(file='pr.worstcasemu.rda',pr)
  file.remove('mu.worstcasemu.rda')
}

if (!file.exists('mu.worstcasemu.rda')) {

## pr.worstcasemu.rda is a huge file with daily precipdata based on ECA&D - generated above
  load('pr.worstcasemu.rda')
```

Then the predictand data is processed: estimate annual mean aggregates and the mean seasonal cycle. Remove stations with large gaps of missing values.

```{r}
  ## Strip away stations with a lot of missing data

## Time series of the annual wet-day freq & mean - for evaluation
## Randomly sub-sample due to excessive volume:
  FW <- annual(pr,FUN='wetfreq',nmin=350)
  MU <- annual(pr,FUN='wetmean',nmin=350)
  fw <- aggregate(pr,month,FUN='wetfreq')
  mu <- aggregate(pr,month,FUN='wetmean')

  n <- apply(pr,2,FUN='nv')
  y1 <- subset(subset(pr,is=(n==max(n))),is=1)
  save(file='mu.worstcasemu.rda',mu,MU,fw,FW,y1,n)
} else load('mu.worstcasemu.rda')

## Keep stations with no missing data and stations
## without suspect outlier trends
nok <- n
ok <- apply(mu,2,function(x) sum(!is.finite(x))==0) &
      apply(fw,2,function(x) sum(!is.finite(x))==0) &
      abs(apply(MU,2,FUN='trend.coef')) <= 1 &
      abs(apply(FW,2,FUN='trend.coef')) <= 0.02
mu <- subset(mu,is=ok) # REB 2015-03-01
MU <- subset(MU,is=ok) # REB 2015-03-01
fw <- subset(fw,is=ok)
FW <- subset(FW,is=ok)

## Remove stations with little data
nval <- apply(MU,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
nval <- apply(FW,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
```

Get and process the predictor data.

```{r}
print("predictor")
if (!file.exists('pre.worstcasemu.rda')) {
  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(-100,30),lat=c(0,40))
  pre <- spatial.avg.field(C.C.eq(t2m))
  attr(pre,'region') <- '100W,30E/0N,40N'
  save(file='pre.worstcasemu.rda',pre)
} else load('pre.worstcasemu.rda')
```

Now the data is ready for the analysis. Calibrate the regression models and extract the regression coeffieients. Use 'apply' to speed up the process for multiple stations.

```{r}
## Extract the monthly aggregates for all stations
print('apply mucal')
V <- apply(mu,2,FUN='mucal',pre=pre)
Beta <- lapply(V,FUN='beta')

## Collect the R-squared statistics from lm(y ~ x) for each site
print('muskill')
r2 <- lapply(V,muskill)
```

**Plot figure 1**

```{r,echo=TRUE}
## Extract the mean, min, max, wettest month, and driest months in terms of mu
## for all stations:
print('muclim')
X <- apply(mu,2,FUN='muclim')
print(table(X[4,]))
wmns <- as.numeric(rownames(table(X[4,])))
nc <- max(wmns) - min(wmns) + 1
cols <- colscal(n=nc)
col1 <- cols[c(X[4,])]
cex <- 1.5*n/max(n)

## The relationship between mu and e_s for one station
## to show the calibration procedure
is <- (1:length(n))[X[4,]==8][1]
print(paste('plot mucal - is=',is))
mucal(subset(mu,is=is),pre=pre,verbose=TRUE,plot=TRUE)

figlab('Figure 1')
#dev.copy2pdf(file='figure1.pdf')
```

**Figure 2a+b**

```{r}
## Remove locations with missing values for PCA 
print('Mu - matrix for PCA')
Mu <- as.matrix(coredata(mu))

## Anomalies wrt the mean value at each location.
Mu <- apply(Mu,2,function(x) (x - mean(x)))
pca <- svd(Mu)

## Plot maps with PCs:
print('Maps with PCs')

mupcamap(mu,pca,1,xlim,ylim,r2)
figlab('Figure 2a')
#dev.copy2pdf(file='figure2a.pdf')
mupcamap(mu,pca,2,xlim,ylim,r2)
figlab('Figure 2b')
#dev.copy2pdf(file='figure2b.pdf')
```

**Figure 3**

```{r,echo=TRUE}
print('Variance accounted for by the modes:')
print(round(100*pca$d**2/sum(pca$d**2),1))

## The sign of PCs is arbitrary...
print(cor.test(pca$v[,1],as.numeric(r2)))
print(cor.test(pca$v[,2],as.numeric(r2)))
print(cor.test(pca$v[,3],as.numeric(r2)))

## Estimate trend statistics for both fw and mu:
## Use subset - in the PCA there is something strange in the 1990s.
fw.trend <- 100*apply(subset(FW,it=c(1950,1990)),2,'trend.coef')/
                      apply(FW,2,'mean',na.rm=TRUE)
mu.trend <- 100*apply(subset(MU,it=c(1950,1990)),2,'trend.coef')/
                      apply(MU,2,'mean',na.rm=TRUE)

#dev.new(); diagnose(MU)

## Select only the sites which have an R2 greater than 0.6:
is <- (1:length(r2))[as.numeric(r2)> 0.6]
mux <- subset(mu,is=is)
MUx <- subset(MU,is=is)
FWx <- subset(MU,is=is)
Xx <- X[,is]
Vx <- V[is]

## Read the GCM ensembles andestimate the 5 and 95 percentiles as well as
## the ensemble mean.

## Get the annual mean temperature from GCM ensembles:
if (!file.exists('cmip5.rda')) {
  print('get CMIP5 RCP4.5')
  rcp4.5 <- readGCMs(path='CMIP5.monthly/rcp45/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP8.5')
  rcp8.5 <- readGCMs(path='CMIP5.monthly/rcp85/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP2.6')
  rcp2.6 <- readGCMs(path='CMIP5.monthly/rcp26/',pattern='tas_Amon_ens_rcp')
  save(file='cmip5.rda',rcp4.5,rcp8.5,rcp2.6)
} else load('cmip5.rda')

## derive time series for each location:
Z.rcp4.5 <- lapply(Vx,'muproject',rcp4.5)
Z.rcp8.5 <- lapply(Vx,'muproject',rcp8.5)
Z.rcp2.6 <- lapply(Vx,'muproject',rcp2.6)
t <- index(rcp4.5)

## Map showing RCP4.5 ensemble mean and the upper 95% change in the
## outer part of the symbol. Also an insert with box-plot diagram
## showin the other RCPs.

## Estimates for mu in 2010:
mu2010.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2010,end=2010)
mu2010.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2010,end=2010)
mu2010.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2010,end=2010)
x2010 <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[3]]))
x2010u <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[2]]))
y2010 <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[3]]))
y2010u <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[2]]))
z2010 <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[3]]))
z2010u <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[2]]))

## Repeat for 2100:
mu2100.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2100,end=2100)
mu2100.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2100,end=2100)
mu2100.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2100,end=2100)
x2100 <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[3]]))
x2100u <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[2]]))
y2100 <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[3]]))
y2100u <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[2]]))
z2100 <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[3]]))
z2100u <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[2]]))

## Data frams with changes in percentages:
mu.2100 <- data.frame(mean.RCP4.5=x2100 - x2010,q95.RCP4.5=x2100u - x2010u,
                      mean.RCP2.6=z2100 - z2010,q95.RCP4.5=z2100u - z2010u,
                      mean.RCP8.5=y2100 - y2010,q95.RCP8.5=y2100u - y2010u)

## Fig 3
## Plot a map of projected values:
#dev.new()
cols <- colscal(n=100,col='precip')
cx2100 <- round(x2100 - x2010)
cx2100u <- round(y2100u - y2010u)
cx2100[cx2100 < 1] <- 1; cx2100[cx2100 > 100] <- 100
cx2100u[cx2100u < 1] <- 1; cx2100u[cx2100u > 100] <- 100
colx <- cols[cx2100]
coly <- cols[cx2100u]
Cex <- 1.25

print("Plot a map of projected values:")
map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',gridlines=FALSE,
    colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
par(xpd=TRUE)
text(20,73,'Wet-day mean: 2100')
legend(20,32,c(expression(bar(x)),expression(q[95])),
       pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)

colbar(pretty(c(x2100u- x2010u, x2100- x2010),n=15),cols,
       fig = c(0.05, 0.1, 0.05, 0.2))

par(new=TRUE,fig = c(0.05, 0.4, 0.75, 0.975),
    cex.axis=0.75,mar=c(1,1,0.1,0.1),xaxt='n')
boxplot(mu.2100,col=c(rep(rgb(0.5,0.5,0.5,0.3),2),
                      rep(rgb(0.5,1,0.5,0.3),2),
                      rep(rgb(1,0.5,0.5,0.3),2)))
par(xaxt='s')
axis(1,c(1,3,5),labels=c('RCP4.5','RCP2.6','RCP8.5'))
grid()

figlab('Figure 3')
#dev.copy2pdf(file='figure3.pdf')
```

**Supporting material/analysis (SM)**

```{r,echo=TRUE}
## Evaluation:
## Use the calibration strategy to predict the annual mu based on the
## predictor (t2m -> e_s)
print('evaluation: correlation')
#mu.eval <- lapply(Vx,'mupredict',annual(pre))
mu.eval <- lapply(Vx,'mupredict',annual(pre),prct=FALSE)
m <- length(mu.eval); n <- length(mu.eval[[1]])
MUz <- matrix(unlist(mu.eval),n,m)
MUz <- zoo(MUz,order.by=index(mu.eval[[1]]))
MUz <- subset(MUz,it=MUx)
MUx <- subset(MUx,it=MUz)
FWx <- subset(FWx,it=MUz)
r.eval <- apply(rbind(coredata(MUz),coredata(MUx)),2,'cormu')
print(summary(r.eval))

## ----------- Supporting Material ------------------------

#dev.new()
par(xaxt='n',yaxt='n',bty='n')
plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
text(0.5,0.5,'Supporting figures',cex=2,font=2)
dev.copy2pdf(file='figureSM0.pdf')

## Plot the statistics of R2:
#dev.new()
hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
     xlab=expression(paste(R^2,' (%)')),freq=TRUE,
     main="Summary of regression scores")
grid()
figlab('Figure SM1')
#dev.copy2pdf(file='figureSM1.pdf')

## Example of estimates for 2050:
print('Example plot - evolution')
#dev.new()
par(bty='n')
i <- 1
N <- length(Z.rcp4.5)
plot(Z.rcp4.5[[i]],plot.type='single',lty=c(2,2,1),lwd=c(1,1,2),
     ylab='%',ylim=c(80,150),
     main=paste('Wet-day mean at',loc(subset(mux,is=i))))
shade(Z.rcp4.5[[i]],col=rgb(0.5,0.5,0.5,0.3))
shade(Z.rcp8.5[[i]],col=rgb(1,0.5,0.5,0.3))
shade(Z.rcp2.6[[i]],col=rgb(0.5,1,0.5,0.3))
grid()
figlab('Figure SM2')
#dev.copy2pdf(file='figureSM2.pdf')

## Trend in projected wet-day mean
trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
## Strange results:
print((1:N)[trendbeta < 0])

## test: Does the model predict observed trends?
print('evaluation: trends')

## Make sure to compare series with data for same times
mask <- !is.finite(coredata(MUx))
class(MUz) <- class(MUx)
muz <- coredata(MUz)
muz[mask] <- NA; dim(muz) <- dim(MUz)
coredata(MUz) <- muz

## Only look at stations with more than 50 years with data
ok <- (apply(coredata(MUx),2,nv) > 50)
MUz <- subset(MUz,is=ok)
MUx <- subset(MUx,is=ok)
FWx <- subset(FWx,is=ok)

#dev.new()
trend.mux <- apply(MUx,2,'trend.coef')
trend.pre <- apply(MUz,2,'trend.coef')
trenderr.mux <- apply(MUx,2,'trend.err')
trenderr.pre <- apply(MUz,2,'trend.err')

## Need to get a picture whether the predictions gives a plausible
## upper limit.
trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
                          y=c(-trend.pre,trend.pre))

xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)

par(bty='n',col.sub='grey',cex.sub=0.8)
plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
     main=expression(paste('Trends in ',mu,': observed and predicted upper limit')),
     xlab='Observed trend (mm/decade)',ylab='predicted trend (mm/decade)',
     xlim=xylim,ylim=xylim,
     sub=paste('Mean correlation for local year-to-year variations over t=[',
       start(MUx),',',end(MUx),
       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
       round(quantile(r.eval,0.95),2),
       ')',sep='') )
grid()

polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(0.2,0.6,1,0.1),border=NA)
polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(1,0.2,0.2,0.1),border=NA)
points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)

## Plot error bars 
apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                     })

figlab('Figure SM3')
#dev.copy2pdf(file='figureSM3.pdf')

## Map showing trends in mu
#dev.new()
map(MU,FUN='trend',colbar=list(breaks=seq(-1,1,length=21),rev=TRUE),cex=cex)
figlab('Figure SM4',ypos=0.999)
figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
#dev.copy2pdf(file='figureSM4.pdf')

## 99-percentile wet-day mean and typical wet-day frequency
## The Tellus paper on specification approximately annual maximum.
## Prob(X>x) for annual maximum for 24hr precip
## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
#dev.new()
par(bty='n')
W <- apply(fw,2,FUN='muclim')
hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
     main='Wet-day percentile for annual maximum 24-precipitation',
     xlab='p (%)')
grid()

figlab('Figure SM5')
dev.copy2pdf(file='figureSM5.pdf')

## Statistics of trend in wet-day frequency
print('Wet-day frequency statistics')
#dev.new()
hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
     main='Trend in wet-day frequency',
     xlab=expression(paste(f[w],' (%/decade)')))
grid()
figlab('Figure SM6')
#dev.copy2pdf(file='figureSM6.pdf')

## Map showing trends in fw
#dev.new()
map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
figlab('Figure SM7',ypos=0.999)
figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
#dev.copy2pdf(file='figureSM7.pdf')

## Figure SM8.
## test: See if the quantiles are consistent when the mean mu varies.
#dev.new()
qtest <- aggregate(y1,year,FUN='qqexp')
qx <- c(coredata(qtest[,1:101]))
qy <- c(coredata(qtest[,102:202]))

par(bty='n')
plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
     main='Test: exponential distribution & changing mean',
     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
lines(c(0,40),c(0,40),col='grey')
figlab('Figure SM8')
#dev.copy2pdf(file='figureSM8.pdf')

pr.mean <- aggregate(y1,by=month,FUN='mean')
pr.mu <- aggregate(y1,by=month,FUN='wetmean')
pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
y1.l <- spell(y1,threshold=1)
pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')

#dev.new()
par(bty='n',xaxt='n')
plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
     col=c('steelblue','darkblue','grey','darkgreen','red'),
     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
grid()
par(yaxt='s',xaxt='s')
axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')

legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1))
figlab('Figure SM9')
#dev.copy2pdf(file='figureSM9.pdf')

## Map of the future return values

print('Map of return values for 2100 with RCP4.5')
rv <- MUx
## x2100u gives the percentage of mu,
coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
dev.new()
map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
figlab('Figure SM10',ypos=0.999)
figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
#dev.copy2pdf(file='figureSM10.pdf')

## Print the numbers:
print('--- Changes in 20% returnvalue in terms of % from 2010:')
print(lapply(mu.2100,summary))
print(paste('Summary for',length(mu.2100[[1]]),'locations'))

## Compare the regression coefficients derived from individual
## seasonal cycle with that derived from mean climatology.

data(mu.eq.f.tx)

col <- rgb(0.1,0.1,0.7,0.25)
mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
b1 <- as.numeric(lapply(Beta,function(x) x[1]))
e1 <- as.numeric(lapply(Beta,function(x) x[2]))

#dev.new()
par(bty='n')
plot(b1,pch=19,col=col,cex=cexr2,
     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
     xlab='',ylab=expression(beta))
grid()
for (i in 1:length(b1)) {
  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
}
polygon(c(1,rep(length(b1),2),rep(1,2)),
        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
figlab('Figure SM12')
#dev.copy2pdf(file='figureSM12.pdf')

## Show the predictor area:
X <- retrieve('data/ncep/air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
figlab('Figure SM13',xpos=0.8,ypos=0.999)
##dev.copy2pdf(file='figureSM13.pdf')
```