### 2016-04-28 CORDEX/ensembles.Rmd
ensemble example for CORDEX

---
title: "ensembles"
author: "Rasmus Benestad"
date: "April 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## **Simulated global mean temperartue**

Combine the simulated global mean temperature with information about the model configurations taken from Table 9-a-1 in the IPCC WG 1 2013 report ("AR5):

```{r}
library('esd')
data("IPCC.AR5.Table.9.A.1")
data("global.t2m.gcm")
gcmnm <- unlist(lapply(attr(global.t2m.gcm$global.t2m.cmip5.rcp45,'meta'),function(x) x$GCM))
n <- length(gcmnm); m <- 12
X <- matrix(rep(NA,n*m),n,m)

IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)
IPCC.AR5.Table.9.A.1$Ocean.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)
```

Index some of the entries which do not contain useful information.

```{r}
iam <- is.element(substr(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,1,8),'Included') |
       is.element(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,'<NA>')
IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[iam] <- IPCC.AR5.Table.9.A.1$Model.Name[iam]
iom <- is.element(IPCC.AR5.Table.9.A.1$Ocean.Component.Name,'Included')
IPCC.AR5.Table.9.A.1$Ocean.Component.Name[iom] <- IPCC.AR5.Table.9.A.1$Model.Name[iom]
```

There are many different similar settings with different names. Try to organise and sort the different types under fewer and simpler categories. Here is a mix of conventions and ways to refer to the resolution.

```{r}
fixagcmnm <- grep('Included',as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name))
IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[fixagcmnm] <- IPCC.AR5.Table.9.A.1$Model.Name[fixagcmnm]
```

```{r}
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(' ','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('-','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(',','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('longitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('latitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Nominally','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('equivalentto','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('FiniteVolume','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Averagedcellsize','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('in','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('and','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('TL','T',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
Tres<- c('T42','T31','T159','T63','T159','T959','T213','T319','N48','N49','C180','C360','R42','N96')
for (ires in Tres) {
  ii <- grep(ires,IPCC.AR5.Table.9.A.1$Horizontal.Grid)
  IPCC.AR5.Table.9.A.1$Horizontal.Grid[ii] <- ires
}
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.factor(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
```

Fix the GCM names:

```{r}
IPCC.AR5.Table.9.A.1$Model.Name <- as.character(IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('-','.',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('_','.',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub(' v1.0','',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('(m)','.m',IPCC.AR5.Table.9.A.1$Model.Name,perl=TRUE,fixed=TRUE)
cesm1 <- grep('CESM1',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name[cesm1] <- 'CESM1'
IPCC.AR5.Table.9.A.1$Model.Name <- as.factor(IPCC.AR5.Table.9.A.1$Model.Name)
gcmnm <- gsub('-','.',gcmnm)
gcmnm <- gsub('_','.',gcmnm)
cesm1 <- grep('CESM1',gcmnm)
gcmnm[cesm1] <- 'CESM1'
cnrm5 <- grep('CNRM.CM5',gcmnm)
gcmnm[cnrm5] <- 'CNRM.CM51'
inmcm4 <- grep('inmcm4',gcmnm)
gcmnm[inmcm4] <- 'INM.CM4'
```

Also need to tidy up some of the description concerning the ocean models

```{r}
IPCC.AR5.Table.9.A.1$z.Co.ord <- as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('-coordinate','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('hybrid','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('sigma-z','sigma',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub(' ','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('Isopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('zisopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- as.factor(IPCC.AR5.Table.9.A.1$z.Co.ord)
```

Make a data.frame that organises the model information for each simulation.

```{r, echo=FALSE}
for (i in 1:n) {
  ii <- is.element(toupper(IPCC.AR5.Table.9.A.1$Model.Name),toupper(gcmnm[i]))
  if (sum(ii)>0) {
    x <- c(as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)[ii][1],
               IPCC.AR5.Table.9.A.1$Number.of.Vert.Levels[ii][1],
               IPCC.AR5.Table.9.A.1$Grid.Top.hPa[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Component.Name.or.type)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Atmos.Chemistry.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Land.Surface.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)[ii][1],
               IPCC.AR5.Table.9.A.1$Number.of.Vertical.Levels[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Ocean.Biogeochemistry.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Sea.Ice.Component.Name)[ii][1])
    print(x)
    X[i,] <- x
  }  else print(paste('no match for ',gcmnm[i]))
}
```

Set the column names

```{r}
X <- as.data.frame(X)
colnames(X) <- c('atm.model','atm.grid',
                 'atm.lev','atm.top',
                 'atm.type','Atm.chem',
                 'Land','oce.model',
                 'oce.lev','oce.z',
                 'oce.bio','Seaice')
X$atm.lev <- as.integer(X$atm.lev)
X$atm.top <- as.numeric(X$atm.top)
X$oce.lev <- as.integer(X$oce.lev)
```

Combine the simulated temperature change with corresponding information about the model configuration

```{r}
dT <- colMeans(window(zoo(global.t2m.gcm[[1]]),start=2070,end=2099))-
      colMeans(window(zoo(global.t2m.gcm[[1]]),start=1970,end=1999))
X <- as.data.frame(cbind(round(dT,2),X))
```

##**Results**

Is the simulated temperature change sensitive to how the atmospheric model is configured?

```{r}
fit.atm <- lm(dT ~ atm.grid + atm.lev + atm.top + oce.lev + oce.z,data=X)
print(summary(fit.atm))
```