### 2016-05-13 CORDEX/pca4predictands.Rmd

---
title: "PCA 4 predictands"
author: "Rasmus Benestad"
date: "May 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demonstration

This R-markdown script provides some of the material presented in <http://www.icrc-cordex2016.org/images/pdf/Programme/Abstracts/Orals/Parallel_B/B3_5_Benestad.pdf> at the ICRC-CORDEX conference <http://www.icrc-cordex2016.org/index.php> in Stockholm, May 2016.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r,warnings=FALSE}
## Script to plot PCA of stations
library(esd)
```

Get the R-package esd (source: <https://github.com/metno/esd>)

```{r}
## Demonstration using real data from the ECA&D: Europe
Y0 <- station(lon=c(-20,20),lat=c(40,60),nmin=70,src='ecad')
map(Y0,FUN='mean')
## Do the analysis for the winter season only (DJF)
Y <- subset(as.4seasons(Y0),it='Jan')

## Need to remove period with little data and stations with little data
Y <- subset(Y,it=c(1950,2015))
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 50)
## Show the data availability
diagnose(Y)

## Fill in missing data
Y <- pcafill(Y)
diagnose(Y)

## compute PCA
pca <- PCA(Y)
plot(pca,new=FALSE)
pca <- subset(pca,pattern=1:4)

## Get a predictor - need to download the file because OpenDap doesn't seem to work
tas <- retrieve('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/air.mon.mean.nc',lon=c(-30,30),lat=c(35,65))
tas <- subset(as.4seasons(tas),it='Jan')
eof <- EOF(tas)

## Example of downscaling
ds <- DS(pca,eof)
plot(ds)
```