### 2016-05-30 Tanzania/dse.tz-ncep.Rmd
Add DSensemble results

---
title: "Tanzania"
author: "Rasmus Benestad"
date: "May 10, 2016"
output: html_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r,message=FALSE}
## Script to downscale temperature from Tanzania.
## Data from Habiba.

library(esd)
```

# Functions

Prepare the predictands

```{r,message=FALSE}
predictands <- function(param='tmin',incl.ghcn=FALSE) {
  ## Prepare the predictand data
  #print("predictands")
  fname <- paste(param,'.tz.rda',sep='')
  path <- paste('Habiba/',param,sep='')
  if (!file.exists(fname)) {
  files <- list.files(path,pattern='rda',full.names=TRUE)
  locs <- list.files(path,pattern='rda')
  dse <- grep('dse.tm',files)
  files <- files[-dse]; locs <- locs[-dse]
  print(files)
  locs <- substr(locs,1,nchar(locs)-4)
  Y <- NULL
  for (i in 1:length(files)) {
    load(files[i])
    y1 <- eval(parse(text=locs[i]))
    if (is.null(Y)) Y <- y1 else
                    Y <- combine.stations(Y,y1)
  }
  save(file=fname,Y)
  plot(Y)
} else load(fname)
  #print('GHCN')
  if (incl.ghcn) {
    xstations <- switch(param,'tmin'='Habiba/tn.eafrica.rda',
                            'tmax'='Habiba/tx.eafrica.rda')
    load(xstations)
    if (param=='tmin') y <- tn.eafrica else y <- tx.eafrica
    y <- subset(y,it=c(1950,2015))
    Y <- combine(Y,y)
  }
  map(Y,FUN='mean',cex=-2)
  invisible(Y)
}
```

Function that applies the downscaling to a combination of predictands in terms of principal component analysis (PCA) results and predictors that involve large multi-model ensembles.

```{r,message=FALSE}
downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
                      lon=c(0,100),lat=c(65,90),eofs=1:6,n=10,rel.cord=FALSE,select=NULL) {

  print('downscale')

## Use a time and space window:  
  Y <- subset(Y,it=period)
  Y <- subset(Y,is=list(lat=lat,lon=lon))

## Estimate seasonal means & weed out stations with little data
  Y4 <- subset(as.4seasons(Y,FUN=FUN,nmin=60),it=it)
  ok <- apply(coredata(Y4),1,nv)
  Y4 <- subset(Y4,it=ok>0)
  nok <- apply(coredata(Y4),2,nv)
  Y4 <- subset(Y4,is=nok>15)

  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',sep=''))
  if (plot) map(Y,FUN=FUN,cex=-2)

  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
  print(paste(nmiss,'% missing',sep=''))

## Fill missing data using PCA-based regression
  Z <- pcafill(Y4)

  pca <- PCA(Z,n=n)
  if (plot) plot(pca)

## Downscale results
  print('DSensemble')
  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
                        biascorrect=TRUE,rcp=rcp,eofs=eofs,select=select,
                        lon=lon,lat=lat,rel.cord=rel.cord,it=it)

  attr(dse.pca,'N.missing') <- nmiss
  invisible(dse.pca)
}
```

Function that extracts results for a given year

```{r}
distill <- function(x,it=2050) {
  y <- subset(x,it=it)
  attributes(y) <- NULL
  y <- mean(y)
  return(y)
}
```

Function that grids the results

```{r,message=FALSE}
gridmap <- function(dse.results,param=NULL,
                    fast=TRUE,it=c(2070,2099),it0=c(1979,2012),verbose=FALSE,
                    breaks=seq(0,7,length.out=15),pal='warm',rev=FALSE) {
  if (verbose) print('gridmap')
  library(LatticeKrig)

  if (is.null(param)) param <- varid(dse.results) 
  if (verbose) print ('as.station')
  if (inherits(dse.results,'pca')) dse.station <- as.station(dse.results)  else
                                   dse.station <- dse.results

  if (verbose) print('distill')
  Z0 <- lapply(dse.station,distill,it=it0)
  Z <- lapply(dse.station,distill,it=it)
  lon <- unlist(lapply(dse.station,lon))
  lat <- unlist(lapply(dse.station,lat))
  alt <- unlist(lapply(dse.station,function(x) alt(attr(x,'station'))))
  X <- attr(dse.station[[1]],'station')
  z0 <- matrix(unlist(Z0),1,length(dse.station))
  z <- matrix(unlist(Z),1,length(dse.station))
  data(etopo5)
  etopo5 <- subset(etopo5,is=list(lon=range(lon),lat=range(lat)))
  etopo5[etopo5<=0] <- NA

  if (verbose) print('grid')
  ## Set the grid to be the same as that of etopo5:
  grid <- structure(list(x=lon(etopo5),y=lat(etopo5)),class='gridList')

  #3 Flag dubplicated stations:
  ok <- !(duplicated(lon) & duplicated(lat))

  ## Spread in the  90-percente interval changing
  obj <- LatticeKrig( x=cbind(lon[ok],lat[ok]),
                      y=z[1,ok] - z0[1,ok],Z=alt[ok] )

  ##  obj <- LatticeKrig( x=cbind(lon[ok],lat[ok]), y=z[2,ok],Z=alt[ok])
  w <- predictSurface(obj, grid.list = grid,Z=etopo5)
  w$z[is.na(etopo5)] <- NA

  ## Get rid of packages that have functions of same name:
  detach("package:LatticeKrig")
  detach("package:fields")
  detach("package:spam")
  detach("package:grid")
  detach("package:maps")

  ## Convert the results from LatticeKrig to esd:
  W <- w$z
  attr(W,'variable') <- param
  attr(W,'unit') <- 'degC' 
  attr(W,'longitude') <- w$x
  attr(W,'latitude') <- w$y
  class(W) <- class(etopo5)

  ## Make a projection that zooms in on the Barents region
  #dev.new()

  colbar <- list(breaks=round(breaks,2),rev=rev,pal=pal)
  if (verbose) print(colbar)
  if (is.null(rev)) rev <- switch(param,'t2m'=FALSE,'pr'=TRUE)
  Wx <- max(abs(W),na.rm=TRUE)
  if (is.null(breaks)) breaks <- round(seq(0,Wx,length=31),2)
  if (fast)
    map(W,xlim=range(lon(W)),ylim=range(lat(W)),
        colbar=colbar,new=FALSE)
  else {
    attr(W,'variable') <- NULL
    attr(W,'unit') <- NULL
    map(W,xlim=range(lon(W)),ylim=range(lat(W)),projection='sphere',
        colbar=colbar,verbose=verbose)
  }
  figlab(paste(it,collapse='-'),ypos=0.99)
}

```

Function that applies tests to theresults through applying PCA

```{r,warning=FALSE}
pcaasses <- function(Y) {
  ## Cut away blocks with missing data:

  nv <- apply(coredata(Y),2,nv)
  Y <- subset(Y,is=nv > 12000)
  Y4s <- as.4seasons(Y,nmin=60)
  y1 <- subset(Y4s,it='Jan')
  y2 <- subset(Y4s,it='Apr')
  y3 <- subset(Y4s,it='Jul')
  y4 <- subset(Y4s,it='Oct')

  ## Fill in missing
  y1 <- pcafill(y1)
  y2 <- pcafill(y2)
  y3 <- pcafill(y3)
  y4 <- pcafill(y4)

  ## Apply PCA:
  pca.djf <- PCA(y1)
  pca.mam <- PCA(y2)
  pca.jja <- PCA(y3)
  pca.son <- PCA(y4)

  ## Examine the results:
  print('DJF - PCA')
  plot(pca.djf,new=FALSE)
  print('MAM - PCA')
  plot(pca.mam,new=FALSE)
  print('JJA - PCA')
  plot(pca.jja,new=FALSE)
  print('SON - PCA')
  plot(pca.son,new=FALSE)

  ## compare the station data with reanalysis - based on CCA
  ## Read reanalysis:
  X <- retrieve('air.mon.mean.nc',lon=range(lon(Y))+c(-3,3),lat=range(lat(Y))+c(-3,3))
  X4s <- as.4seasons(X)

  eof.djf <- EOF(subset(X4s,it='Jan'))
  eof.mam <- EOF(subset(X4s,it='Apr'))
  eof.jja <- EOF(subset(X4s,it='Jul'))
  eof.son <- EOF(subset(X4s,it='Oct'))

  cca.djf <- CCA(pca.djf,eof.djf)
  cca.mam <- CCA(pca.mam,eof.mam)
  cca.jja <- CCA(pca.jja,eof.jja)
  cca.son <- CCA(pca.son,eof.son)

  print('DJF - CCA')
  plot(cca.djf,new=FALSE)
  print('MAM - CCA')
  plot(cca.mam,new=FALSE)
  print('JJA - CCA')
  plot(cca.jja,new=FALSE)
  print('SON - CCA')
  plot(cca.son,new=FALSE)

## The quick and effiecint way to downscale all the stations:
  ds.djf <- DS(subset(pca.djf,pattern=1:4),eof.djf)
  ds.mam <- DS(subset(pca.mam,pattern=1:4),eof.mam)
  ds.jja <- DS(subset(pca.jja,pattern=1:4),eof.jja)
  ds.son <- DS(subset(pca.son,pattern=1:4),eof.son)

  print('DJF - DS')
  plot(ds.djf,new=FALSE)
  #dev.new()
  print('MAM - DS')
  plot(ds.mam,new=FALSE)
  #dev.new()
  print('JJA - DS')
  plot(ds.jja,new=FALSE)
  #dev.new()
  print('SON - DS')
  plot(ds.son,new=FALSE)

  return(list(pca.djf=pca.djf,pca.mam=pca.mam,pca.jja=pca.jja,
              pca.son=pca.son,y.djf=y1,y.mam=y2,y.jja=y3,y.son=y4))
}
```

#Maximum temperature

The code below shows how the functions are used and applied to the data for the daily maximum $T_x$. The first chunck of code carries out some preliminary analysis, the second shows how the results are derived, and the third show how the plots are made.

```{r,message=FALSE,warning=FALSE}
#-----------------------------------------------------------------------
# Define season and parameter

param <- 'tmax'
FUN <- 'mean'
reanalysis <- 'air.mon.mean.nc'
#reanalysis <- 'ERAINT_t2m_mon.nc'
FUNX <- 'mean'
data.path <- 'dse.Tz.NCEP/'

predictands(param) -> Y

Y <- subset(Y,it=c(1961,2011))
lon=range(lon(Y))+c(-10,10); lat=range(lat(Y))+c(-15,15)

## Check the predictands:
X <- pcaasses(Y)
```

The PCA suggest that most of the variance in the station data for daily maximum temperature $T_x$ can be represented by a small number of leading modes. For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.

The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor. The closest spatial similarity is seen in SON.

The DS analysis applied to the reanalysis suggests strongest match in MAM and weakest in DJF. The spatial patterns differ to some extent and there is not a great match between the predictor and predictand.

```{r,message=FALSE,warning=FALSE}
## Get the predictand -> Y

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  #print(paste('Generating dse.',param,'.tz.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''))) {
    #dev.new()
    dse.t2m.tz.rcp45 <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''),dse.t2m.tz.rcp45)
  }

  #print(paste('Generating dse.',param,'.tz.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp26 <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''),dse.t2m.tz.rcp26)
  }

   #print(paste('Generating dse.',param,'.tz.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp85 <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''),dse.t2m.tz.rcp85)
  }
}
#print('--- Completed downscaling ---')
```

##The March-May season.

```{r,message=FALSE,warning=FALSE}
## analysis:

load(paste(data.path,'dse.',param,'.tz.rcp85.mam.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4g.',param,'.pdf'))

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4h.',param,'.pdf'))

gridmap(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.pdf',sep=''))

validate(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.diagnose.pdf',sep=''))
```

The plume plots look OK for Arusha and Musoma, and the observations fit reasonably well within the envelope of downscaled ensemble (RCP8.5). The gridded map suggests strongest warming in the north (~5C increase in 2099).

The rank test (Wilcox) indicates reasonable scores, although the rank distribution is not entirely flat. The diagnostic (target plot) for the diagnosis suggest that the trend in the ensemble is higher than in observation (horizontal axis) and that the inter-annual variance (vertical axis) is smaller in the ESD results.

##The December-February season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.djf.rda',sep=''))  
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.pdf',sep=''))
validate(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.validate.pdf',sep=''))
diagnose(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.diagnose.pdf',sep=''))
```

A weaker warming (~2C) is projected in DJF, also with strongest response in the north.  But the rank scores are weak.

##The June-August season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.jja.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.pdf',sep=''))
validate(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.validate.pdf',sep=''))
diagnose(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.diagnose.pdf',sep=''))
```

The assessment suggest the JJA season is better downscaled, with reasonable rank scores. The projected warming is more uniform over Tanzania, albeit with some local variations.

##The September-November season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

```

The SON season is associated with strongest warming in the north, and displays a clear north-south gradient. The warming is projected to be 0.7--1.3C by 2099, and the rank scores are moderately good - the distribution of the ranks had a U-shape rather than being flat.

##Check a random station: is is well-represented by the PCA?

```{r,message=FALSE,warning=FALSE}
## Check the PCA-results:
z.djf <- as.station(X$pca.djf)
y.djf <- X$y.djf
par(new=FALSE)
plot.zoo(subset(z.djf,is=1),subset(y.djf,is=1),
         main=paste('Test: Winter temperature at',loc(subset(y.djf,is=1))))
plot(subset(z.djf,is=1),lwd=3,col='grey70')
lines(subset(y.djf,is=1))
```

The single station data could be well-recovered from the PCA product.

#Implementation - Minimum temperature

Get and prepare the predictand data for the daily minimum temperature $T_n$.

```{r,message=FALSE,warning=FALSE}
#-----------------------------------------------------------------------
# Define season and parameter

param <- 'tmin'

predictands(param) -> Y

Y <- subset(Y,it=c(1961,2011))
lon=range(lon(Y))+c(-10,10); lat=range(lat(Y))+c(-15,15)

## Check the predictands:
X <- pcaasses(Y)
```

Apply the ESD models to implement thedownscaling

```{r,message=FALSE,warning=FALSE}
## Get the predictand -> Y

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  print(paste('Generating dse.',param,'.tz.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''))) {
    #dev.new()
    dse.t2m.tz.rcp45 <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''),dse.t2m.tz.rcp45)
  }

  print(paste('Generating dse.',param,'.tz.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp26 <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''),dse.t2m.tz.rcp26)
  }

   print(paste('Generating dse.',param,'.tz.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp85 <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''),dse.t2m.tz.rcp85)
  }
}
print('--- Completed downscaling ---')
```

The PCA suggest that most of the variance in the station data for daily maximum temperature $T_n$ can be represented by a small number of leading modes. For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.

The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor.

The DS analysis applied to the reanalysis suggests strongest match in MAM and weakest in DJF. The spatial patterns differ to some extent and there is not a great match between the predictor and predictand.

##The March-May season

```{r,message=FALSE,warning=FALSE}
## analysis:

load(paste(data.path,'dse.',param,'.tz.rcp85.mam.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4g.',param,'.pdf'))

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4h.',param,'.pdf'))

gridmap(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.pdf',sep=''))

validate(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.validate.pdf',sep=''))

diagnose(z,new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.diagnose.pdf',sep=''))
```

The plume plots look good for Arusha and Musoma, and strongest warming (3.5-4C) is projected for the eastern part of Tanzania. The evaluation of the quality gave moderate results - some discrepancies in terms of the rank scores.

##The December-February season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.djf.rda',sep=''))  
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.pdf',sep=''))
validate(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.diagnose.pdf',sep=''))
```

Strongest warming projected for the southeast part of the country (~3C), and the evaluation suggest moderate scores. Again, still some discrepancies.

##The June-August season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.jja.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.pdf',sep=''))
validate(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.diagnose.pdf',sep=''))
```

Most pronounced warming is projected for the souther part (~3C), and the assessment indicates moderate quality with some discrepancies. The ensemble results tend to account for lower inter-annual variability than in the observations and weaker trend (except for one station with higher trend).

##The September-November season

```{r,message=FALSE,warning=FALSE}
#load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
load(paste(data.path,'dse.',param,'.tz.rcp45.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('SON')

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('SON')
```

Strongest projected warming in the south (~3C) and moderate quality.

## Check a random station: is is well-represented by the PCA?

```{r,message=FALSE,warning=FALSE}
## Check the PCA-results:
z.djf <- as.station(X$pca.djf)
y.djf <- X$y.djf
plot.zoo(subset(z.djf,is=1),subset(y.djf,is=1),
         main=paste('Test: Winter temperature at',loc(subset(y.djf,is=1))))
plot(subset(z.djf,is=1),lwd=3,col='grey70')
lines(subset(y.djf,is=1))
```

Single station records can be recovered from the PCA.

#Summary

There are problems with some stations, which is visible in the PCA, CCA and DS carried out with just reanalysis. The spatial patterns indicate non-smooth and incoherent structures, suggesting problems with some of the stations and a mismatch with the reanalysis. Furthermore, the cross-validation reveal moderate scores (r typically less than 0.7), and the evaluations based on trends, range of inter-annual variability and rank score reveal some discrepancies.

The NCEP/NCAR reanalysis is probably better than ERAINT for ESD model calibration.