### 2016-06-13 CixPAG/CixPAG-aphrodite.Rmd

---
title: "CixPAG-aphrodites"
author: "Rasmus Benestad"
date: "May 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
library(esd)
```

## Including Plots

```{r}
odurl <- 'http://iridl.ldeo.columbia.edu/SOURCES/.RIHN/.aphrodite/.V1101/.APHRO_MA_050deg/.precip/dods'
con <- nc_open(url)
print.ncdf4(con)
```