### 2016-07-29 gmst-outlook/gmst-outlook.Rmd
update

---
title: "gmst-outlook"
author: "Rasmus Benestad"
date: "July 28, 2016"
output: pdf_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r}
library(esd)
download.file('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/air.mon.mean.nc',
              destfile = 'air.mon.mean.nc')
```

Get the NCEP/NCAR reanalysis 1 from FTP and estimate the global mean anomaly

```{r}
t2m <- retrieve('air.mon.mean.nc')
gmst <- aggregate.area(t2m,FUN='mean')
gmsta <- anomaly(gmst,ref=1961:1990)
```

Find the most recent complete month and set up a regression model

```{r}
t <- index(gmst)
mon <- month(t[length(t)])
Y <- annual(gmsta)
x <- aggregate(subset(gmsta,it=1:mon),year,FUN='mean')
z <- lm(coredata(Y) ~ coredata(x))
y <- predict(z,newdata=x)
n <- length(y)
```

Plot the analysis

```{r}
ylim<- range(c(coredata(Y),coredata(y)),na.rm=TRUE) + 0.25*c(-1,1)
plot(zoo(x),zoo(Y),ylim=ylim,
     xlab=paste('Jan-',month.abb[mon],' mean',sep=''),
     ylab='annual mean', sub='base line: 1961-1990; source NCEP/NCAR reanalysis 1',
     main='Global mean temperature',pch=19,cex=1.5)
lines(rep(coredata(x)[n],2),ylim,lwd=7,col='grey90')
arrows(coredata(x)[n],coredata(y)[n],min(x,na.rm=TRUE),coredata(y)[n],lty=2)
points(coredata(x)[n],coredata(y)[n],cex=1.5,lwd=2)
text(coredata(x)[n],min(y),max(year(x)),sub=1)
grid()
```