### 2016-10-17 CixPAG/CixPAG-imd.Rmd
Something new

---
title: "CixPAG"
author: "Rasmus Benestad"
date: "August 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Reading data predictand

```{r pressure, echo=FALSE}
pt <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
map(tx)
```

```{r}
Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
eof.tx <- EOF(Tx.maxjja)

Tx.nhd <- annual(tx,FUN='count',threshold=40)
```