### 2016-10-18 KSS/ds-t2m4KSS.Rmd
update from VGL

---
title: "Downscale t2m for KSS"
author: "Rasmus Benestad"
date: "October 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Function definitions:

```{r}
downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
                      lon=c(-20,40),lat=c(50,80),ip=1:6,n=6,
                      rel.cord=FALSE,select=NULL) {

  print('downscale')

## Use a time and space window:  
  Y <- subset(Y,it=period)
  Y <- subset(Y,is=list(lat=lat,lon=lon))

## Estimate seasonal means & weed out stations with little data
  Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
  ok <- apply(coredata(Y4),1,nv)
  Y4 <- subset(Y4,it=ok>0)
  nok <- apply(coredata(Y4),2,nv)
  Y4 <- subset(Y4,is=nok>15)

  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',
  sep=''))
  if (plot) map(Y,FUN=FUN,cex=-2)

  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
  #print(paste(nmiss,'% missing',sep=''))

## Fill missing data using PCA-based regression
  Z <- pcafill(Y4)

  pca <- PCA(Z,n=n)
  if (plot) plot(pca)

## Downscale results
  print('DSensemble')
  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
                        biascorrect=TRUE,rcp=rcp,ip=ip,select=select,
                        lon=lon,lat=lat,rel.cord=rel.cord,it=it)

  attr(dse.pca,'N.missing') <- nmiss
  invisible(dse.pca)
}
```

## Retrieving the data

Get the predictand data: daily temperature from ECAD:

```{r predictand-data}
if (!file.exists('t2m.nordic.rda')) {
  ss <- select.station(param='tg',lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
  Y <- station(ss)
  save(Y,file='t2m.nordic.rda')
} else load('t2m.nordic.rda')
```

```{r}
## Only use the most recent 66 years 
Y <- subset(Y,it=c(1960,2015))
## Remove records with many missing datapoints
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 14000)
```

```{r}
#-----------------------------------------------------------------------
# Define season and parameter

FUN <- 'mean'
reanalysis <- 'air.mon.mean.nc'
lon=c(-30,30); lat=c(50,75)
FUNX <- 'mean'
verbose=TRUE

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  print(paste('Generating dse.kss.t2m.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
    save(file=paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''),Z)
  }

  print(paste('Generating dse.kss.t2m.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste('dse.kss.t2m.rcp26.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
    save(file=paste('dse.kss.t2m.rcp26.',it,'.rda',sep=''),Z)
  }

   print(paste('Generating dse.kss.t2m.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste('dse.kss.t2m.rcp85.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
    save(file=paste('dse.kss.t2m.rcp85.',it,'.rda',sep=''),Z)
  }
}
print('--- Completed downscaling ---')
```

## Post-processing

```{r}
## source the functions for gridding
demo(pca2eof,ask=FALSE)
load('dse.kss.t2m.rcp45.djf.rda')
## Filter and reduce data size by reteining only the 4 leading PCA
Z <- subset(Z,ip=1:4)
## Convert the PCA object to EOF 
Z <- as.eof.dsensemble.pca(Z)
save(Z,file='dse.KSS.t2m.rcp45.djf.rda')
map(Z,it=c(2040,2060))
```