### 2016-10-19 CixPAG/CixPAG-imd.Rmd

---
title: "CixPAG"
author: "Rasmus Benestad"
date: "August 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Reading data predictand: maximum temperature

```{r IMD maxmum temperature}
tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
map(tx)
```

Some initial exploration of the data - the mean summer maximum daily maximum temperature

```{r}
Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
eof.tx <- EOF(Tx.maxjja)
plot(eof.tx,new=FALSE)
```

Trend in the men summer maximum daily maximum temperature

```{r}
map(Tx.maxjja,FUN='trend',new=FALSE)
```

The mean number of days with daily maximum temperature exceeding 40C.

```{r}
Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
map(Tx.nhd40,new=FALSE)
```

## Precipitation

```{r IMD maxmum temperature}
rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
```

Monsoon rain statistics for the past - any interesting trends?

```{r}
mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')
MU <- annual(rr,FUN='wetmean')  ## Use capital letters for theentire year
fw <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetfreq')
FW <- annual(rr,FUN='wetfreq')
pt <- annual(rr,FUN='sum')
map(pt,new=FALSE)
```

The precipitation intencity during the monsoon months

```{r}
map(mu,new=FALSE)
map(mu,FUN='trend',new=FALSE)
```

The precipitation frequency during the monsoon months

```{r}
map(fw,new=FALSE)
map(fw,FUN='trend',new=FALSE)
```