### 2016-10-20 CixPAG/CixPAG-imd-precip.Rmd
minor fixes

---
title: "CixPAG-IMD precip"
author: "Rasmus Benestad"
date: "October 20, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

## Precipitation

```{r IMD_precip}
rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
```

Monsoon rain statistics for the past - any interesting trends?

```{r aggregated_precip}
mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')
MU <- annual(rr,FUN='wetmean')  ## Use capital letters for theentire year
fw <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetfreq')
FW <- annual(rr,FUN='wetfreq')
pt <- annual(rr,FUN='sum')
map(pt,new=FALSE)
```

The precipitation intencity during the monsoon months

```{r}
map(mu,new=FALSE)
map(mu,FUN='trend',new=FALSE)
```

The precipitation frequency during the monsoon months

```{r}
map(fw,new=FALSE)
map(fw,FUN='trend',new=FALSE)
```

## Downscaling

```{r max_summer_temps}
mu.jjas <- annual(subset(rr,it=month.abb[6:9]),nmin=4)
index(mu.jjas) <- year(mu.jjas)
eof.Y <- EOF(mu.jjas,n=5)
plot(eof.Y,new=FALSE)
```

```{r}
predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
predictor <- maskland(predictor,land=TRUE)
predictor <- C.C.eq(predictor)
predictor <- annual(subset(predictor,it='jja'),nmin=3)
eof.X <- EOF(predictor,n=10)
plot(eof.X,new=FALSE)
```

Check the link between large and small scales

```{r}
class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
ds.test <- DS(eof.Y,eof.X,ip=1:6)
plot(ds.test,new=FALSE)
```