### 2016-11-28 CixPAG/CixPAG-spell.Rmd

---
title: "CixPAG - spells"
author: "Rasmus Benestad"
date: "November 3, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Scatter plots: mean versus standard deviation of anomalies. 
Combine several stations to link mean with mean spell length?

Downscaling of heat wave spell length. Paper. Wheat.