### 2016-12-07 EU-Circle/eu-circle_case-study-SL.Rmd

---
title: "Sea level"
author: "Rasmus Benestad"
date: "December 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Evidence for coherent the local sea level rise based on PCA. Relate the local response to the global, expecting a physical connection

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
library(RgoogleMaps)
```

## Analysis

You can also embed plots, for example:

```{r sea-level-data}
gsl <- GSL()
gloss <- station.gcloss()
sonel <- station.sonel()
newlin <- station.newlin()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.