### 2017-04-06 EU-Circle/eu-circle_case-study-5.Rmd

---
title: "EU-Circle Case study 5"
author: "Rasmus Benestad"
date: "May 2, 2016"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rapid Winter Flooding (melting ice, narrow mountain streams, flooding) around Dresden, Germany

Dresden is the largest city in the Eastern part of Germany, Saxony, near the Czech border. It is crossed by the large river Elbe (its width is around 110m in Dresden) which comes from the Czech Republic and flows through Magdeburg and Hamburg into the North Sea. The region between Dresden and the Czech border, but also in the near surroundings in the Eastern and South-Western directions, is quite hilly with mountains up to ca. 1200 m. In the recent past (especially notable are the big floodings in 2002 and 2013) there has been quite some floodings caused by intense and long rain which led to floods of the Elbe.However, there is another threat which is happening more often and hence causing more damage because of the currently ongoing climate changes. In the mountains are lots of small mountain streams which are really small on a normal day (a typical one has a width of 1.5m and a depth of 0.5m). But in cases of heavy rains these small streams rapidly grow quite large (1-2h is not uncommon), especially compared to their normal size (sometimes they grow to their hundredfold size in respect of the amount of water they carry). This causes a lot of damage, especially as prediction and short-term prevention mechanisms are not really possible because of the small timescale. The damages caused by such events are quite local and often not related to the flooding of big rivers. In these  regions there are often roads going through valley which are vital in a sense that if they are  not available any more, quite long bypasses have to be taken. Additionally railways,  especially railway bridges, are often effected as well. These two aspects show the impacts on the transport network.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

## Analysis: precipitation

```{r, eval=FALSE}
## Get files on the outgoing long-wave radiation (OLR).
cmip5.urls <- function(experiment='rcp45',varid='rlut',
                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
  urlfiles <- "NA"
  if(verbose) print("cmip5.urls")
  if(is.null(path)) path <- getwd()
  for (iexp in experiment) {
    if(verbose) print(iexp)
    for (ivar in varid) {
      if(verbose) print(ivar)
      ## Loop on the number of experiments
      for (irun in 0:110) { ## 
        if(verbose) print(paste(irun))
        ## Update experiment number
        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
        else run.id <- as.character(irun)

        ## Create output directory for the climate experiment
        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
                              fsep = .Platform$file.sep)
        if (!file.exists(path.exp)) dir.create(path.exp)
        if (verbose) print(path.exp[grep(iexp,experiment)])
        ## Define the output file
        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
        destfile <- paste(destfile,"_Amon_ens_",sep="")
        destfile <- paste(destfile,iexp,sep="")
        destfile <- paste(destfile,run.id,sep="_")
        destfile <- paste(destfile,".nc",sep="")

        if (!file.exists(destfile) | force) {
          ## Update output filename with attributes:
          urlfile  <- paste(url,ivar,sep="")             # add var directory
          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
        }
        urlfiles <- c(urlfiles,urlfile)
        if (verbose) print(urlfile)
      }
    } # End for   
  }
  return(urlfiles[-1])
}

## download the CMIP5-files from the KNMI ClimateExplorer
urls <- cmip5.urls()
for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))

```

Data from ECA&D: precipitation. Snow depth?

```{r}
library(esd)
ss <- select.station(param='precip',cntr='Germany',alt=250,nmin=75)
map(ss,new=FALSE)
Y <- station(ss)
Y <- subset(Y,it=c(1950,2015))
diagnose(Y)
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 24000)
```

```{r}
y <- subset(Y,is=1)
coredata(y) <- rowMeans(coredata(Y))
attr(y,'location') <- 'German higher elevation'
attr(y,'longitude') <- mean(lon(Y))
attr(y,'latitude') <- mean(lat(Y))
attr(y,'altitude') <- min(alt(Y))
plot(y)
lines(trend(y),lty=2)
grid()
```

Area total precipitation from gridded EOBS.

```{r}
rr <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=range(lon(Y)),lat=range(lat(Y)))
map(rr,new=FALSE)
Arr <- aggregate.area(rr)
plot(rr,new=FALSE)
```

Wet day mean and frequency

```{r}
mu <- annual(rr,FUN='wetmean')
fw <- annual(rr,FUN='wetfreq')
MU <- aggregate.area(mu,FUN='mean')
FW <- aggregate.area(fw,FUN='mean')
plot(MU,new=FALSE)
lines(trend(MU),lty=2)
grid()
plot(FW,new=FALSE)
grid()
lines(trend(FW),lty=2)
```

Plot the probability of more than 50mm/day

```{r}
Pr.50 <- FW*exp(-50/MU)
plot(Pr.50)
lines(trend(Pr.50),lty=2)
grid()
```

Test to see if the long-wave outgoing radiation (OLR) can be used as a predictor for the winter-time wet-day mean precipitation over Germany. It contains information about cloud tops, and the hypothesis is that heavy precipitation comes from clouds with high (and cold) cloud tops.

```{r, fig.width=8, fig.height=8}
olr <- retrieve('~/Downloads/ERAINT-OLR-mon.nc', lon=c(-50,40),lat=c(40,70))
olr <- subset(as.4seasons(olr),it='djf')
eof.olr<- EOF(olr,n=5)
class(eof.olr)[3] <- "season"
mu.djf <- subset(as.4seasons(Y,FUN='wetmean'),it='djf')
pca.mu <- PCA(pcafill(mu.djf),n=5)
ds.mu.orl <- DS(pca.mu,eof.olr)
plot(ds.mu.orl,new=FALSE)
```

The calibration is fairly successful in identifying a connection between OLR and precipitation intencity. To see if other factors such as moisture has an effect, we estimate the residual from this regression analysis and repeat the analysis using saturation water vapour as predictor.

```{r, fig.width=8, fig.height=8}
## Estimate the residuals
pca.res <- subset(pca.mu,it=ds.mu.orl)
coredata(pca.res) <- coredata(pca.res) - coredata(ds.mu.orl)
plot(pca.res)
```

```{r, fig.width=8, fig.height=8}
mat <- retrieve('~/Downloads/ERAINT-t2m-mon.nc', lon=c(-90,40),lat=c(30,70))
es <- C.C.eq(mat)
es <- subset(as.4seasons(es),it='djf')
es <- mask(es,land=TRUE)
eof.es<- EOF(es,n=10)
class(eof.es)[3] <- "season"
ds.mu.es <- DS(pca.res,eof.es)
plot(ds.mu.es,new=FALSE)
```

A projection of future wet-day mean precipitation based on DSensemble will be carried out in two steps. First use the PCA from the original data as predictands in a set of runs with OLR as predictors. Then use the residual as estimated above as predictand for a set of DSensembles runs with the water vapour pressure as predictors. The final result is the sum of these two steps.The wet-day mean precipitation is projected through s single stemp where DSensemble uses SLP as predictor. The seasonal means are then estimated as the product between the wet-day frequency and the wet-day mean precipitation.

## Analysis: temperature

Check for melting.

```{r}
ss <- select.station(param='tn',cntr='Germany',alt=500,nmin=75)
map(ss)
X <- subset(station(ss),it=month.abb[1:4])
diagnose(X)
X <- subset(X,it=c(1940,2015))
```

```{r}
tnx <- annual(X,FUN='max',nmin=100)
plot(tnx,new=FALSE)
grid()
```