### 2017-09-15 MIMOZA/mimoza.Rmd
mimoza + Bangladesh

---
title: "MIMOZA"
author: "Rasmus Benestad"
date: "September 14, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r esd}
library(esd)
```

## Tidal data and sea level

You can also embed plots, for example: <http://www.psmsl.org/data/obtaining/>,

```{r sealevel, echo=FALSE}
sl.maputo <- read.table('http://www.psmsl.org/data/obtaining/met.monthly.data/986.metdata',sep=';')
sl.island <- read.table('http://www.psmsl.org/data/obtaining/met.monthly.data/1054.metdata',sep=';')
```

## IMILAST storm tracks

```{r}
if (!file.exists('mimoza.imilast.rda')) {
  imilast.storms <- read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt',lon=c(30,90),lat=c(-50,-5))
  cyclones <- events2trajectory(imilast.storms)
  save(cyclones,file='mimoza.imilast.rda')
} load('mimoza.imilast.rda')
```

## GHCND temperature and precipitation

```{r}
## get the GHCND data
if (!file.exists('mimoza.ghcnd.rda')) {
  ## Daily maximum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmax')
  tmax <- station(ss,verbose=TRUE)
  ## Daily minimum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmin')
  tmin <- station(ss,verbose=TRUE)
  ## Daily precipitation
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
  precip <- station(ss,verbose=TRUE)
  ## Set bad (missing) data to NA
  z <- coredata(tmax); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmax)
  z <- coredata(tmin); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmin)
  z <- coredata(precip); z[z < 0] <- NA; z[z > 700] <- NA; z -> coredata(precip)
  ## Save in R-binary for convenience.
  save(tmax,tmin,precip,file='mimoza.ghcnd.rda')
} else load ('mimoza.ghcnd.rda')
```