### 2017-09-20 R3/R3-WP4/R3_ESD_NVE.Rmd
R3 folder added into esd_Rmarkdown

Worshop on ESD 4 NVE 
-----------------------------------
title: "esd4nve_examples"
author: "Abdelkader Mezghani"
date: "March 11, 2016"
output: html_document
---

This is a generated R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

---------------------------------

## **Data & Structure**

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/DATA/data_esd_2016/fwddataforr3workshopenom2uker")
```

## Require the following libraries
```{r required R packages,echo=FALSE}
require(esd)
require(leaflet)
require(sp)
require(rgdal)
```

## R3 ESD & NVE
### Function for reading discharge data into the esd format

```{r readNVE,warning=FALSE}
## Create a function that converts NVE format to esd format
readNVE <- function(file='/data_esd_2016/atnsjo.csv',loc='Atnsjo',lat='61',lon='10',
                    alt=1204,stid = '2.32',plot=TRUE,...) {

  data <- read.csv(file=file,header=FALSE, comment.char="#", 
                   dec=".", sep=" ", stringsAsFactors=FALSE)

  dates <- as.Date(data$V1,format="%Y%m%d")
  xz <- zoo(data$V2,order.by=dates)

  y <- as.station(x=xz,loc=loc,lon=lon,lat=lat,alt=alt,stid=stid,
                  param="Q",
                  unit="m/s^{-2}",
                  longname="Discharge data",...)
  y[y==-9999] <- NA ## Replace missing values with NA	
  invisible(y)
}
```

#### Read meta data from the shape file

```{r}
library(rgdal)
library(leaflet)

rsf <- readOGR(dsn = 'Shape_filer_R3_WP4/R3_ESD_nbf_gruppe1.shp',layer = 'R3_ESD_nbf_gruppe1', GDAL1_integer64_policy = TRUE)
rsf.lonlat <- spTransform(rsf,CRSobj = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

str(rsf)

df <- data.frame(rsf@data, stringsAsFactors = FALSE)

xy.coords <- as.data.frame(t(apply(as.matrix(1:18), 1, function(i) unlist(rsf@polygons[[i]]@labpt))))
colnames(xy.coords) <- c('x','y')
coordinates(xy.coords) <- ~ x + y
#CRS(xy.coords) <- rsf@proj4string
proj4string(xy.coords) <- rsf@proj4string

lonlat.coords <- spTransform(xy.coords,CRSobj = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

lonlat <- attr(unclass(lonlat.coords),'coords')

meta <- data.frame(station_id = as.character(df$STASJON_NR), location = as.character(df$ST_NAVN), longitude = lonlat[1,] ,latitude = lonlat[,2])

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = 'orange')

leaflet(rsf.lonlat) %>% 
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addPolygons(data = rsf.lonlat,stroke = TRUE, weight = 1) %>%
  addAwesomeMarkers(lng = lonlat[,1] , lat = lonlat[,2],icon = icons)

```

#### Reading data into esd format
```{r,warning=FALSE} 
library(esd)

if (!file.exists('R3_WP4_Qdata_sep2017/q.nve.rda')) {
  lf <- list.files('R3_WP4_Qdata_sep2017',pattern = '.csv',full.names = TRUE) ## remember to update the file path name
  readName <- function(lf) {
    namefromfile <- function(i) unlist(strsplit(basename(lf),split = '_')[[i]][2])
    fname <- apply(as.matrix(1:length(lf)),1,namefromfile)
    invisible(substr(fname,1,nchar(fname) - 4))
  }

  locs <- tolower(rsf$ST_NAVN)
  staID <- substr(rsf$STASJON_NR,1,nchar(as.character(rsf$STASJON_NR))-2)
  ## Display list of discharge stations 
  listStationName <- readName(lf)
  for (i in 1 : length(lf)) {
    id <- grep(as.character(staID[i]), tolower(basename(lf)))
    ifile <- lf[id]
    cat(i, tolower(locs[i]),ifile,sep = '\n')
    #stopifnot(length(id) == 1)
    x <- readNVE(file= ifile,loc=locs[i],lon=lonlat[i,1],lat=lonlat[i,2], stid = staID[i], alt=NA , cntr='norway')
    if (i ==1) X <- x  else X <- combine.stations(X,x)
  }

  # Visual ckeck of the results
  plot(X,plot.type = 'mult',ylab = toupper(substr(locs,1,5)),cex.lab = 0.8,new = FALSE)

  # save to file 
  q.nve <- X 
  save(q.nve,file = 'R3_WP4_Qdata_sep2017/q.nve.rda')
} else 
  load('R3_WP4_Qdata_sep2017/q.nve.rda')
```

#### Q Data Analysis
```{r data Analysis, warning=FALSE,eval=TRUE, eval=TRUE, warning=FALSE}
setwd('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/')
load('q.nve.rda')
library(leaflet)

content <- paste(sep = "<br/>", paste("<b><a>",loc(q.nve),"</a></b>"))

# plot(climatology(as.monthly(q.nve)),new=FALSE) ## Plot the climatology/hydrograph

# compute annual max
Qall.ann.max <- annual(q.nve,FUN="max")
plot(Qall.ann.max,plot.type = 'mult')
grid()

# compute annual sum
Qall.ann.sum <- annual(q.nve,FUN="sum")
plot(Qall.ann.sum,plot.type = 'mult')
plot(Qall.ann.sum,plot.type = 'mult')

# perform a quick trend analysis
Qall.tr.ann.max <- apply(Qall.ann.max,2,FUN='trend.coef',na.rm=TRUE)
Qall.mean.ann.max <- apply(Qall.ann.max,2,FUN='mean',na.rm=TRUE)
tr.rel <- round(Qall.tr.ann.max / Qall.mean.ann.max *100,digits = 1) ## in % per decade

# test significancy
Qall.pval.ann.max <- round(apply(Qall.ann.max,2,FUN='trend.pval',na.rm=TRUE),digits = 2)
q.nve.sig <- subset(q.nve, is = which(Qall.pval.ann.max <= 0.05))

# Construct the map
content <- paste(sep = "<br/>", paste("<b><a>",toupper(loc(q.nve)),"</a></b>"), paste(as.character(tr.rel)," % "))
pal <- colorBin(rev(colscal(20,'t2m')),c(-10,10),bins = 8, pretty = TRUE)
df <- data.frame(lon = lon(q.nve),lat = lat(q.nve), trend = as.numeric(tr.rel))

leaflet() %>% 
  addProviderTiles(providers$Esri.WorldStreetMap,
                   options = providerTileOptions(noWrap = TRUE)) %>%  
  addCircleMarkers(lng  = df$lon, lat = df$lat, radius = 8, fillOpacity = 0.8,
                   stroke = TRUE, color = 'black', weight = 1, fill = TRUE, fillColor = pal(df$trend), popup = content) %>%
  addCircleMarkers(lng  = lon(q.nve.sig), lat = lat(q.nve.sig), radius = 2, fillOpacity = 0.8, stroke = TRUE, color = 'black', weight = 1, fill = TRUE, fillColor = 'black') %>%
  addLegend(position = 'bottomleft',pal = pal, values = pal(df$trend), 
            title = 'Trends in Q [%]')

```

## Downscaling monthly totals single discharge data @ Atnasjø

```{r,warning=FALSE,eval= FALSE}
library(esd)
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
selQ <- subset(q.nve,is = 1)
```

## Search the nearest meteorological station 
```{r}
# select nearest preciptiation stations
path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'

ssP <- select.station(src='metnod',param=c('precip'))

if (!file.exists(file.path(path.data,'selNearPre.rda'))) {
  mdis <- distAB(lon(selQ),lat(selQ),ssP$longitude,ssP$latitude)
  # 20 closest stations
  selP <- subset(as.data.frame(ssP), subset = is.element(order(mdis),c(1:20)))
  PRE <- station.metnod(stid = selP$station_id,param='precip',user = 'metno')
  save(PRE,file = file.path(path.data,'selNearPre.rda'))
} else {
  load(file.path(path.data,'selNearPre.rda'))
}

# Pgood <- clean.station(P,miss = 0.1,verbose = FALSE)
# plot(Pgood)
# title(paste(loc(Pgood)))
```

### Select nearest temperature station
```{r selNearT}
path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'

ssT <- select.station(src='metnod',param=c('t2m'))

if (!file.exists(file.path(path.data,'selNearT2m.rda'))) {
  # select nearest temperature station
  mdis <- distAB(lon(selQ),lat(selQ),ssT$longitude,ssT$latitude)
  selT <- subset(as.data.frame(ssT), subset = is.element(order(mdis),c(1:20))) # top 20 stations
  T2m <- station.metnod(stid = selT$station_id, param = 't2m',user = 'metno')
  plot(T2m,plot.type = 'mult')
  Tgood <- subset(T2m, is = 16)
  #Tgood <- clean.station(T2m,miss = 0.1,verbose = FALSE)
  plot(Tgood)
  # title(paste(loc(Tgood)))
  save(T2m, file = file.path(path,'selNearT2m.rda'))
} else {
  load(file.path(path.data,'selNearT2m.rda'))
}
```

### Produce map with all layers
```{r selNearP}
labq <- paste(sep = "<br/>", paste("<b><a>",loc(q.nve),"</a></b>"))
labP <- paste(sep = "<br/>", paste("<b><a>", ssP$location,"</a></b>"))
labT <- paste(sep = "<br/>", paste("<b><a>", ssT$location,"</a></b>"))

leaflet() %>% 
  addProviderTiles(providers$Esri.WorldStreetMap,group = 'streetmap') %>%
  addPolygons(data = rsf.lonlat,stroke = TRUE, weight = 1 , group = 'NVE catchments') %>%
  setView(lng = 15, lat = 65, zoom = 4) %>%
  addMarkers(lng = lon(q.nve),lat = lat(q.nve), icon = icons, group = 'Discharge (red)',
             popup = labq) %>%
  #addAwesomeMarkers(lng = lon(q.nve),lat = lat(q.nve), icon = icons, group = 'Nearest') %>%
  addCircleMarkers(lng = ssT$longitude,lat = ssT$latitude, stroke = TRUE, weight = 1, 
                   radius= 3, color = 'orange', popup = labT, group = 'Thermometers (orange circles)') %>%
  addCircleMarkers(lng = ssP$longitude,lat = ssP$latitude, stroke = TRUE, weight = 1, 
                   radius= 3, color = 'blue', popup = labP, group = 'Raingauges (blue circles)') %>%
  addLayersControl(overlayGroups = c("Discharge (red)","Raingauges (blue circles)","Thermometers (orange circles)","NVE catchments"), options = layersControlOptions(collapsed = FALSE))
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.4seasons(precip.ERAINT(lon = c(0,30), lat = c(50,85))),it = 'djf') 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.4seasons(q1),it = 'djf') # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.annual(subset(precip.ERAINT(lon = c(0,30), lat = c(50,85)),it = 'djf'),nmin=0),it = c(1979,2013)) 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.annual(subset(q1,it = 'djf'),nmin=0),it=c(1979,2013)) # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.annual(subset(t2m.ERAINT(lon = c(0,30), lat = c(50,85)),it = 'djf'),nmin=0),it = c(1979,2013)) 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.annual(subset(q1,it = 'djf'),nmin=0),it=c(1979,2013)) # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r monthly flow prediction, eval = TRUE}
# find and extract common dates

# set the common time Period
commonPeriod <- c("1979-01-01","2002-03-31")
t2m <- subset(Tgood, it = commonPeriod)
precip <- subset(Pgood, it = commonPeriod)
q <- subset(selQ, it = commonPeriod)

# Perform a correlation map with large scale features.
T2M <- t2m.ERAINT(lon = c(0,30), lat = c(50,85))
map(T2M)
T2M <- subset(T2M,it = commonPeriod)
corT <- corfield(as.monthly(q),T2M)

PRE <- precip.ERAINT(lon = c(0,30), lat = c(50,85))
PRe <- subset(PRE, it = commonPeriod)
PRe.jan <- subset(PRe, it = 'jan')
q.jan <- subset(as.monthly(q), it = 'jan')
corP <- corfield(q.jan, PRe.jan,verbose = TRUE)

# plot the time series
Y <- combine.stations(precip,t2m,q)
plot(Y, plot.type = 'mult')

P <- as.zoo(precip)
t2m <- as.zoo(t2m)
u <- as.zoo(q)

par(bty="n")
plot(merge(P,u,as.anomaly(t2m)),plot.type='single',col=c("steelblue","black","red"),lwd=c(3,1,2))
dev2bitmap('hydroex-daily.png',res=150)

Pmm <- aggregate(P,as.yearmon,FUN='mean')
Tmm <- aggregate(t2m,as.yearmon,FUN='mean')
umm <- aggregate(subset(u, it = commonPeriod),as.yearmon,FUN='mean')

Pmm <- nearest.field(PRe, is = umm)
plot(Pmm)
Tmm <- nearest.field(T2m, is = umm)
plot(Tmm)

#plot(coredata(P),coredata(u))

cal <- data.frame(u=coredata(umm),t2m=coredata(Tmm),Pmm=coredata(Pmm))
fit <- lm(u ~ t2m + Pmm,data=cal)
print(summary(fit))
plot(coredata(umm),predict(fit,newdata = cal))
cor(coredata(umm),predict(fit,newdata = cal), use = 'complete.obs')
par(bty="n")
plot(coredata(umm),predict(fit,newdata = cal),
     main='Monthly river flow predicted from monthly T and P',
     ylim=range(coredata(Pmm)),
     xlab=expression(u[hydro]),ylab=expression(u==f(T[2*m],P)))
lines(range(coredata(Pmm)),range(coredata(Pmm)),col="red",lty=2)
grid()
dev2bitmap(file='hydroex.png',res=150)

temp <- as.station(t2m,lon=lon,lat=lat,alt=alt,unit='deg C',param='t2m',loc=loc,stid=stid)

if (FALSE) {
  trace(z.u <- try(DSensemble.t2m(umm, path = '~', predictor='data/ERA40/ERA40_precip_mon.nc', biascorrect=TRUE,plot=TRUE)))
  dev2bitmap('hydroex-dse-z.t2m.png',res=150)
  save(file='hydroex-dse-z.t2m.rda',z.t2m)
}

temp <- Tmm
if (FALSE) {
  trace(z.t2m <- try(DSensemble.t2m(temp, path = '~', predictor='~/ERA40_t2m_mon.nc', biascorrect=TRUE,plot=TRUE)))
  dev2bitmap('hydroex-dse-z.t2m.png',res=150)
  save(file='hydroex-dse-z.t2m.rda',z.t2m)
}

precip <- as.station(P,lon=lon,lat=lat,alt=alt,unit='mm/day',param='precip',loc=loc,stid=stid)
if (FALSE) {
  z.pre <- try(DSensemble.precip(precip,biascorrect=TRUE,plot=TRUE))
  dev2bitmap('hydroex-dse-z.pre.png',res=150)
  save(file='hydroex-dse-z.pre.rda',z.pre)
}

flow <- as.station(u,lon=lon,lat=lat,alt=alt,unit='m^3/s',param='u',loc=loc,stid=stid)

muam <- annual(precip,FUN='exceedance',threshold=1)
fwam <- annual(precip,FUN='wetfreq',threshold=1)
uamx <- annual(flow,FUN='count',threshold=30)
attr(uamx,'variable') <- 'n(X > m^3/s)'
attr(uamx,'unit') <- ' '

am <- data.frame(u=coredata(uamx),mu=coredata(muam),fw=coredata(fwam))
emu <- glm(u ~ mu + fw, data=am, family='poisson')

plot(uamx,lwd=3,ylim=c(0,70))
lines(zoo(exp(predict(emu)),order.by=index(uamx)))

text(1980,70,pos=4,'Number of annual events: u > 30 m³/s')
legend(1980,65,c("hydrologic model",expression(f(mu,f[w]))),
       lty=1,col=c("red","black"),lwd=c(3,1),bty="n")
dev2bitmap('hydroex-u.gt.30.png',res=150)

```

```{r}
PRE <- precip.ERAINT(lon = c(0,30),lat =c(50,80))
Q <- subset(q.nve,is = 1)

ds.Q <- DS(subset(annual(Q),it = c(1950,2015)), EOF(subset(annual(PRE),it=c(1950,2015))))
plot(ds.Q)
```
```{r}
cor(subset(annual(Q),it=c(1960,2010)), subset(annual(subset(PRE,is=1)),it = c(1960,2010)),use = 'complete.obs')
```