### 2017-12-13 R3/R3-WP4/RCubed_wrk_hbe.Rmd
updated Rmd-file testing things and added a file with command lists to hack on wrf-files to makethem nice for esd

---
title: "RCubedwrk"
author: "Helene B. Erlandsen"
date: "November 16, 2017"
output:
  pdf_document: default
  html_document: default
---

##GCM Files
downloaded from Norstore using python script .... and a modified version of *wps.py* (provided by MP, UiB) to be compatible with Python 3.6.3 (a change in char byte-char handeling needed specification of utf8-format)

##RCM Files
9 years and one month of dynamically downscaled bias corrected NorESM-data have been downloaded and processed using cdo and nco (getRCMfiles.py) to provide files compatible with ESD retrieve.rcm. Note that the calender is noleap or 365day. The files are from the 3-hourly wrfxtrm-output, and has a res of about 4 km covering South Norway.

The processing includes extracting variables t2m,pr,u10,v10,q2,tskin,precip and aggregating to daily and montly means. For all variables but precip the monthly standard deviation is also stored.

For the dynamically downscalled bias corrected NorESM WRF run some cdo nco hacking is done to give the files 2-D latitude and longitude variables and dimensions x,y, and CF-time (cdo does not like wrf). The same is done to the WRF file with the geograhical features (geo_em). The cdo and nco commands are stored in scripts saved for later.

The daily precipitation is converted to mm per day, to allow aggregation to monthly wet-day mean and wet day frequence with a wet-day threshold of 1 mm/day. Remember weight by noleap month if aggregating from month to season.

The (nc4) RCM files:
wrf_sub3hr_NorESM_BC_2000_2010.nc 16GB compressed,
wrf_day_NorESM_BC_2000_2010.nc 2GB maybe comressed, do not remember
wrf_mon_NorESM_BC_2000_2010.nc 228M uncompressed

They are at lustre under A under my username

## Data Reading & Structure (copied parts from Abdelkader)

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE) 
```

### Require the following libraries
Also added a boolean for update esd (provided that a new version is pulled)
```{r required R packages,echo=FALSE}
needed<-c("ncdf4", "zoo")
update_esd <- FALSE
for (package in needed) { 
  if (!require(package,character.only = TRUE, quietly = TRUE)) { 
      install.packages(package);library(package,character.only = TRUE)
    }
}

if ((!require('esd',character.only = TRUE, quietly = TRUE))| (update_esd) ){ 
      install.packages("~/git/esd",repos=NULL,type="source");library('esd',character.only = TRUE)
    }
#require(esd)  
#require(leaflet)
#require(sp)
#require(rgdal)
```

```{r setpath}
#
library(esd)
#source('~/git/esd/R/retrieve.R')
rcmdatapath <- '~/storah/RCubed/data/RCM'
rcmgeopath <-'~/storah/RCubed/data/RCM/WRF_filetypes'
gcmdatapath <- '~/storah/RCubed/data/GCMraw'
gcmbcdatapath <- '~/storah/RCubed/data/GCMBC'
eraidatapath<-'~/storah/RCubed/data/ERAI'
RCMgeo<-'R3geo_emd03run1.nc'
RCMmonp <- 'wrf_mon_NorESM_BC_2000_2010.nc' 
GCMBCmon <- 'NorESM_BC4R3_esd_2000_2010.nc'
GCMmon <- 'NorESM_nonBC4R3_esd_2000_2010.nc'
ERApl <- 'EraI_pl_4_esd_mdfa_197901to201706.nc' 
ERAsf <- 'EraI_sfc_4_esd_mdfa_197901to201706.nc'

```

```{r, warning=FALSE}
gridn <- par(mfrow=c(4, 5))
RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = FALSE)
# just testing
attr(RCMt2m,'greenwich')=TRUE
RCM_oro <- retrieve.rcm(ncfile='geo_em_d03ED.nc',path=rcmgeopath,param='HGT_M',verbose = TRUE)

#map.station(RCM_oro,FUN ="mean",cex=0.3,new=F, is=list(lon=c(3.75,15),lat=c(58,64)))

attributes(RCM_oro) <- NULL
attr(RCMt2m,'altitude') <- RCM_oro

RCMt2m <- subset(RCMt2m, is=list(lon=c(3.75,15),lat=c(58,64)))
map.station(RCMt2m,FUN ="mean",cex=0.3,new=F)
par(gridn)

#check later the stripes

```

```{r,warning=FALSE}
RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = FALSE)
RCMfw <- subset(RCMfw, is=list(lon=c(3.75,15),lat=c(58,64)))
gridn <- par(mfrow=c(4, 5))
# just testing
map.station(RCMfw,FUN ="mean")
par(gridn)
```

```{r loadGMCBC,warning=FALSE,fig.height=32}

GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = FALSE)
attr(GCMBCt2m,'Calendar') <- 'noleap'
GCMBCt2m<-subset(GCMBCt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))

map.field(GCMBCt2m,FUN ="mean", new=F)

```

```{r loadGMCnonBC,warning=FALSE}
#loading non BC data
GCMt2m <- retrieve(ncfile=GCMmon, path=gcmdatapath,param='tas',verbose = FALSE)
attr(GCMt2m,'Calendar') <- 'noleap'
GCMt2m<-subset(GCMt2m,is=list(c(-10,15),c(45,75)),it=c("2000-12-01","2010-12-01"))

#map.field(GCMt2m,FUN ="mean", new=F)

```

```{r loadERAI,warning=FALSE}

ERAIt2m <- retrieve(ncfile=ERAsf, path=eraidatapath,param='t2m',verbose =FALSE)
ERAIt2m<-subset(ERAIt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
gridn <- par(mfrow=c(4, 5))
map.field(ERAIt2m,FUN ="mean")
par(gridn)
```

```{r corrfield test, warning=FALSE,eval=FALSE}
# the station object : pca
# the field object : eof
# later check out https://github.com/metno/esd_Rmarkdown/blob/master/KSS/dse4KSS.R
# 4 similar examples, there Y is station values
# LTR try DS on EOF RCM predictand on EOF GCM predictor
# bc EOFs contain pattern of variance biases in different GCMs not very important, and can try same model for all benestad or try leave on out as in Walton
# latr try common eof append [Re-analysis or GCM-BC | GCM raw] -->time  to find common eofs
#need to consider that we have a short time-series. maybe need monthly data 12*9, monthly anomalies, or seasonal 4*9

#Regridding and making seasonal instead of monthly to look at EOF
RCMt2m.regrid <- regrid.station(RCMt2m,is=list(lon=seq(4,15,by=0.1),lat=seq(58,64,by=0.25)),verbose=T)
# OBS if loaded demos The following object is masked from ‘package:esd’: unit, that is why unit in regrid.station does not work
RCMt2m.seasons.regrid <- as.4seasons(RCMt2m.regrid)
# Make EOFs of the RCM 
eof.test <- EOF(RCMt2m.seasons.regrid)
plot(eof.test)
# Make EOF of the seasonal anaomaly (monthly?)
eof.test <- EOF(as.anomaly(RCMt2m.seasons.regrid))
plot(eof.test)
plot(EOF(GCMBCt2m))

eof.gcm <- EOF(as.anomaly(GCMBCt2m))
plot(eof.gcm)
plot(eof.gcm,ip=2)
ds <- DS(eof.test,eof.gcm)
plot(ds) #r=0.79
plot(ds,ip=2)
plot(eof.test,ip=2)

#Slightly better result from BC model (which is boundary for dyn-DS)
#eof.gcmnoBC <- EOF(as.anomaly(GCMt2m))
#dsnoBC <- DS(eof.test,eof.gcmnoBC)
#plot(dsnoBC) #r=0.76

#Also trying ERA-I  really bad bc not source of dyn RCM
eof.eraI <- EOF(as.anomaly(ERAIt2m))
#dsEra <- DS(eof.test,eof.eraI)
#plot(dsEra) #r=0.3
#dev.new(); plot.zoo(subset(pca.y,ip=1:4))
#dev.new(); plot.zoo(subset(pca.y,ip=1:4))

RCMt2m.seasons.eof <- EOF(subset(RCMt2m.seasons.regrid,it='djf'),n=3)
plot(RCMt2m.seasons.eof,ip=1)

```

```{r pca_on_rcm_seasonal_anomalies}
#Making seasonal and looking at PCA

# Walton method was PCA on each of 60 warming patterns (not warming pattern here) with regional means removed
RCMt2m.ses <-as.4seasons(as.monthly(RCMt2m))
RCMt2m.ses.ano <-as.anomaly(RCMt2m.ses)
RCMt2m.ses.anomaly.pca <- PCA(RCMt2m.ses.ano,n=3)
plot.pca(RCMt2m.ses.anomaly.pca, new=F)
plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=2)
plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=3)
attr(RCMt2m.ses.anomaly.pca,'greenwich')=TRUE
#RCMtm.clim <- as.climatology(as.monthly(RCMt2m))
```

To do the converison of PCA to EOF we need to load some extra packages not loaded by defualt into *esd*. We call the demo-scripts for *esd* *gridmap* and *pca2eof* to load the packge *LatticeKrig*, *spam*, *grid*, *maps*, *fields* [*grid::unit* will replace the *esd*-function *unit*]

```{r loadingaux, include=FALSE}

#This this is based on a package from NCAR using kringing taking altitude into account, so need wrf-oro
# since oro is lacking now making dummy oro

#pca2eof uses etop5 min is around 8-10 km, lower than original grid of 4 km about does it matter?
reload_auxpckg=F
if (reload_auxpckg==TRUE){
  demo(package='esd');demo(gridmap,ask=FALSE);demo(pca2eof,ask=FALSE)
}

```

```{r rcmpca2eof_cca_gcm_eof}
RCMt2m.seasons.eof.from.pca <- pca2eof(RCMt2m.ses.anomaly.pca) 
#Compare not same bca regridding and dummy oro
#dev.new()
plot(RCMt2m.seasons.eof.from.pca)
#dev.new()
plot(RCMt2m.ses.anomaly.pca)

#vis(RCMt2m.ses.anomaly.pca)
#Compare
#dev.new()
#plot(RCMt2m.seasons.eof,ip=2)
#dev.new()
#plot(RCMt2m.seasons.pca,ip=2)

eof.gcm <- EOF(as.anomaly(as.4seasons(GCMBCt2m)))
cca <- CCA(RCMt2m.ses.anomaly.pca,eof.gcm)
plot(cca,new=T) #RCMt2m.pca <- PCA(RCMt2m)

```

```{r test combine EI noresm}
#know from before that ERAI has poor poor temporal covariance as expected with DS norESM
comb <- combine(as.anomaly(as.4seasons(ERAIt2m)),as.anomaly(as.4seasons(GCMBCt2m)))
ceof <- EOF(comb, n=5)
plot(ceof,new=T)
ds <- DS(RCMt2m.seasons.eof.from.pca,ceof)
plot(ds, new=T)

```